### 2019-05-28 Analysis/Simulation/report.Rmd

---
title: "JMP report"
author: "Lucas Avezum"
date: "May 14, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)                                                             ## excel files and,
library(readxl)                                                            ## load the libraries.
library(data.table)
library(foreign)
library(lubridate)
library(lfe)
library(stargazer)
library(ineq)
library(DescTools)
library(gglorenz)

```
```{r, include=FALSE}
load("../../Data/Temp/Raw_data_2.Rda")
lc.wholesale <- data %>% filter(Portfolio_lv1 == "Wholesale") 
lc.wholesale <- Lc(lc.wholesale$PD,lc.wholesale$EAD*lc.wholesale$LGD)

lc.retail <- data %>% filter(Portfolio_lv1 == "Retail") 
lc.retail <- Lc(lc.retail$PD,lc.retail$EAD*lc.retail$LGD)
```

- Average Capital savings using $\hat{P}$ and LGD original distribution: 32.8%

- Average Gini coefficient $PD$ weighted by $LGD$: 0.67

- Correlation between Gini coefficient and capital savings: 0.79

- $\hat{P}=\frac{\sum_i^nEAD_iLGD_iPD_i}{\sum_i^nEAD_iLGD_i}$ \newline

```{r lorenz curve, echo=FALSE,}
plot(lc.wholesale,
     col="black",
     type="l",      # !is not working
     lty=1,
     lwd=3,
     main="Lorenz Curve by Portfolio"     
)

lines(lc.retail, lty=2, lwd=3)

legend("topleft",
       c("Wholesale", "Retail"),
       lty=c(1,2,3),
       lwd=3)
```