### 2019-08-19 Analysis/Draft/Draft.Rmd
First version of paper structure with to-do list

---
title: "Credit risk distribution and capital savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn}
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
anonymous: no
fontsize: 12pt
geometry: margin= 1in
keywords: Key1, Key2, Key3
linkcolor: blue
biblio-style: apsr
subtitle: ''
thanks: Very preliminary draft
#abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
#  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
#  a concave mapping from probabilities of default to capital requiremets. As a consequence,
#  banks can reduce requirements or increase risk-taking by widening their portfolio
#  credit risk distribution.'
toc: no
urlcolor: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

rm(list = ls())

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables

p_w <- c(0.01, 0.99)         ## Set winsorizing limits

```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

```{r change data, eval= FALSE, include=FALSE}
# Expand Pillar-III data frame to plot densities and lorenz curves
expanded.list <- list()
for(i in c("EAD","EL")) {
  expanded.list[[i]] <- pillar3.data %>%
  filter(Method2 %in% c("IRB")) %>%
  select(name,Country,Method,Portfolio_1,EAD,LGD,PD,year)%>%
  mutate(EL  = round(EAD*LGD/1000),
         EAD = round(EAD/1000))%>%
    na.omit(cols=c(i))%>%
    expandRows(i) %>%
    mutate(freq_type = as.factor(i))
}
pillar3.expanded <-  rbindlist(expanded.list, use.names=TRUE,fill = TRUE)

# Aggregate Pillar-III data by Portfolio
portfolio.data <- pillar3.data %>% 
  group_by(Country,Bank,year,Portfolio_1,Portfolio_2) %>%
  mutate(weight=EAD*LGD)%>%
  filter(weight>0, Portfolio_1 %in% c("Wholesale","Retail"))%>%
  dplyr::summarize(gini             = DescTools::Gini(PD,weight),
                   K_calculated     = sum(K_calculated),
                   K_counterfactual = sum(K_counterfactual)) %>%
  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)

bank.data <- bank.data %>%
group_by(bvdid) %>% 
  mutate(savings_log  = log(K_counterfactual-K_calculated),
         Total.assets = Total.assets/10^6,
         Total.Capital= Total.Capital/10^6,
         Tier.1.Capital=Tier.1.Capital/10^6,
         asset_log    = log(Total.assets),
         capital_log  = log(Total.Capital),
         tier.1_log   = log(Tier.1.Capital),
         savings_rate = (K_counterfactual-K_calculated)/K_calculated,
         savings_alt  = (K_counterfactual-K_original)/K_original,
         error        = (K_calculated - K_original)/K_original,
         asset_d      = Winsorize(log(Total.assets)-lag(log(Total.assets),order_by=year), 
                                  probs = p_w ,na.rm = TRUE),
         capital_d    = Winsorize(log(Total.Capital)-lag(log(Total.Capital),order_by=year), 
                                  probs = p_w ,na.rm = TRUE),
         PD_mean_d    = Winsorize(PD_mean-lag(PD_mean,order_by=year), 
                                  probs = p_w ,na.rm = TRUE),
         savings_d    = Winsorize(savings_log-lag(savings_log,order_by=year), 
                                  probs = p_w ,na.rm = TRUE),
         car_d  = Winsorize(Total.Capital.Ratio/100-lag(Total.Capital.Ratio/100,order_by=year), 
                            probs = p_w ,na.rm = TRUE),
         roe_d  = Winsorize(ROE.using.P.L.before.tax..../100-lag(ROE.using.P.L.before.tax..../100,order_by=year),
                            probs = p_w ,na.rm = TRUE)) %>%
  ungroup()

test <- cross.section.decomposition %>% mutate(mean.total.gain = mean(total.gain,na.rm = TRUE)) %>%
  mutate_at(vars(SA.gain:total.gain),  funs((. - mean.total.gain)^2))%>%
summarize_at(vars(SA.gain:total.gain),  sum, na.rm =  TRUE)%>%
  mutate_all( funs((. / total.gain)))
```

# To-Do List {-}

\begin{itemize}

\item Data collection

   \begin{itemize}

   \item[$-$] Collect information for 32 IRB banks that participate on EBA exercise out of 52. Currently, we have information for 10 banks (and 35 SA banks).

   \end{itemize}

\item Write data section:

   \begin{itemize}

   \item[$-$] Use decomposition to show how the intensity and extensity of PD distribution effect varies in time and across banks.

   \item[$-$] Use decomposition to calculate how much of the PD distribution effect variation explains total RW variation (both in time and across banks).

   \item[$-$] Explain capital savings measure

   \item[$-$] Summary statistics

   \end{itemize}

\item Analysis

   \begin{itemize}

   \item[$-$] EBA vs non-EBA:Implement matching strategy following \cite{gropp2018banks}.

   \item[$-$] Dif-on-Dif for EBA sample. Use capital shortfall as treatment intensity indicator.

   \item[$-$] OLS relating capital savings measure to, capital ratio, RWA, ROA, ROE, Z-score, leverage.

   \end{itemize}

\item Robustness
   \begin{itemize}

   \item[$-$] Previous analysis with alternative capital savings measures (Gini coefficient, RW)

   \end{itemize}

\item Extension

   \begin{itemize}

   \item[$-$] Use Barth measures to study interaction with regulation.

   \item[$-$] Use corporate governance \citep{anginer2016corporate} and subordinated debt \citep{gropp2018banks} to study different incentive to raise capital.

   \item[$-$] Use NPL ratios to check if PD dispersion is associated with better risk management practices [@cucinelli2018credit].

   \item[$-$] Check portfolio-level effects, for instance, corporate vs real estate.

   \end{itemize}
\end{itemize}
\newpage

# Introduction

# Background

In this section I describe the chosen method to measure differences in risk distributions:

\begin{equation}
\label{eq1}
\text{RWA}= 12.5 \sum_{i}EAD_iLGD_i f(PD_i),
\end{equation}

\begin{equation}
\label{eq2}
\widehat{\text{RWA}}= 12.5 f(\overline{PD})\sum_{i}EAD_iLGD_i,
\end{equation}

where $i$ is one asset of the portfolio, $f()$ is the regulatory formula and,

\begin{equation*}
\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
\end{equation*}

The capital savings measure is the difference between equations \ref{eq1} and \ref{eq2}:

\begin{equation}
\label{eq3}
\text{RWA savings}=\widehat{\text{RWA}}-\text{RWA},
\end{equation}

\newpage

# Data

```{r}
savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation \\ref{eq3}. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, log.RWA.s, FUN = median, na.rm=TRUE),
                          y = log.RWA.s,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings (USD mn)") +
  theme_bw()

```

\newpage

## Capital Ratio Savings

\begin{equation}
\label{eq4}
\text{Counterfactual capital ratio}=\frac{\text{Capital}}{\widehat{\text{RWA}}},
\end{equation}

\begin{equation}
\label{eq5}
\text{Capital ratio savings}=\text{Counterfactual capital ratio}- \text{Observed capital ratio},
\end{equation}

```{r}
savings_cap = "The figure plots the capital ratio savings measure. This measure is calculated using equation \\ref{eq5}. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, CAR.dif, FUN = median, na.rm=TRUE),
                          y = CAR.dif,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital ratio savings (%)") +
  theme_bw()

```

```{r}
tier1_cap = "The figure shows scatter plots and the linear relationship between the capital savings measure and capital ratios. The y-axis variable on plots a and b is the tier 1 capital ratio, and on plots c and d is the total capital ratio. Plots a and c plots all data points in the sample. Plots b and d plots the average across time for bank using the IRB approach."
```

```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
 plot1 <- bank.data %>%  ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

```{r}
fig3_cap = "PD distribution by bank. The figure shows smoothed PD distributions for each bank in our data. Banks are ranked by median PD and colored by host-country"
```

```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
pillar3.expanded %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 0.05) +
  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    x = 'Probability of default') +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())
```

\newpage

## Decomposition

Rearranging \ref{eq3} and dividing by EAD we have:

\begin{equation}
\begin{aligned}
RW & =  \widehat{\text{RW}} +\text{RWA savings} \\
   & =  \widehat{\text{RW}} +\text{RW}^s 
\end{aligned}
\end{equation}

I use the method proposed by @cannata2012inside to decompose the components of risk-weights.

\begin{equation}
\begin{aligned}
\Delta \text{RW} & = q_{B}^{SA} \Delta \text{RW}^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \widehat{\text{RW}}^{j} \\ 
& + (1-q_{B}^{SA})\sum_{j=1}^{P} \widehat{\text{RW}}^{j} \Delta q^{j}  + (\text{RW}^{SA}-\widehat{\text{RW}})\Delta q^{SA} -(1-q_{B}^{SA})\Delta \text{RW}^{S}+ \text{RW}^{S}\Delta q^{SA}\\
& = \text{Other effects} -\underbrace{(1-q_{B}^{SA})\Delta \text{RW}^{S}}_\text{Intensive margin}+ \underbrace{\text{RW}^{S}\Delta q^{SA}}_\text{Extensive margin}
\end{aligned}
\end{equation}

```{r}
time.dec_cap = "Time decompostion. The figure shows the contribution of each component compared to the time of IRB introduction (t=0). The figure and calculations include only banks with observations for at least 5 consecutive years. The black line is the net change. Year decompostion. The figure shows the contribution of each component compared to 2008. The figure and calculations include all banks in the sample. The black line is the net change." 
```

```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}

plot1 <-time.decomposition %>% 
  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain")),
         RW = RW - t0.RW,
         time = as.factor(time)) %>%
  ggplot(mapping = aes(x = time,
                       y    = value, 
                       fill = condition)) +
geom_bar(stat = 'identity')+
geom_line(aes(group=1, y = RW))  +
  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
 theme(legend.position = c(0.3, 0.1), 
        legend.title = element_blank(), 
        legend.background=element_blank())

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year)) 
plot2 <- year.decomposition %>% 
  select(year,RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
theme(legend.position='none')
grid.arrange(plot1, plot2, ncol=2)

```

```{r}
cross.dec_cap = "Cross-section decompostion. The figure shows the contribution of each component compared to the average bank. The figure and calculations include all banks in the sample. The black line is the total RW. Relative cross-section decompostion. The figure shows the relative contribution of each component compared to the average bank. The figure and calculations include all banks in the sample. " 
```

```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(order.s = abs(int.gain))%>%
  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = reorder(name, total),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  geom_line(aes(y = total,group=1))  +
  labs(x=NULL, y= NULL,subtitle = "Plot a") +
    coord_flip() +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
 theme(legend.position='none')

plot2 <- cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(order.s = abs(int.gain))%>%
  select(name, Country,total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(value = abs(value),
         condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = reorder(name, total),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity',position = "fill")+
  labs(x=NULL, y= NULL,subtitle = "Plot b") +
    coord_flip() +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
   theme(legend.title = element_blank(), 
        legend.background=element_blank(),
       legend.text=element_text(size=7),
       axis.text.y = element_blank())

grid.arrange(plot1, plot2, ncol=2)

```

\newpage

```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  subset(select = c("log.RWA.s","tier1ratio","tier.hat","tier.dif","totalcapitalratio","CAR.hat","CAR.dif","ROE","ROA","Zscore","leverage")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Log RWA savings","Tier 1 ratio","Counterfactual tier 1 ratio","Tier 1 ratio savings","Total ratio","Counterfactual total ratio","Capital ratio savings","ROE","ROA","Z-score","Leverage ratio"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

```{r}
fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE. The solid line is the average across banks for each year. The first and third quartiles are shown in the gray area."
```

```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth", eval=FALSE, include=FALSE}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = TRUE),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

plot1 <- bank.data %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu) %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>% 
ggplot(aes(x=time)) +
 geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Capital savings') +
  theme_bw()

plot2 <- bank.data  %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu) %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'ROE') +
  theme_bw()

plot3 <- bank.data  %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Tier 1 ratio') +
  theme_bw()

grid.arrange(plot1, plot2, plot3, ncol=3)
```

# Results

```{r}

pre.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2009","2010"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

post.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2012","2013"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

DiD.data <- bank.data %>%
  filter(year %in% c("2010")) %>%
  mutate(Shortfall.RW.2 = ifelse(EBAcountry == 1,
                                 ifelse(EBAbank == 1,Shortfall.RW, 0), NA))%>%
  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
    filter(EBAcountry == 1,
           name != "UniCredit",
           name != "Bank of Cyprus") %>%
  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
         Shortfall.mn.2 = ifelse(EBAbank == 1,Shortfall.mn, 0),
         Shortfall.RW.2 = ifelse(EBAbank == 1,Shortfall.RW, 0),
         out.RWA.total  = log.RWA.total.post - log.RWA.total.pre,
         out.RWA.other  = log.RWA.other.post - log.RWA.other.pre,
         out.RWA.credit = log.RWA.credit.post - log.RWA.credit.pre,
         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
         out.K          = log.K.post - log.K.pre,
         out.tier       = log.tier.post - log.tier.pre,
         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
         out.tier.dif   = tier.dif.post - tier.dif.pre,
         Country        = as.factor(Country)) %>% 
  ungroup() %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

```
```{r, include=FALSE}
ggplot(DiD.data, aes(x= EBAbank, y =  out.tier.ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r, include=FALSE}

reg.1 <- felm(out.RWA.s ~ Shortfall.mn.2 | 0 |0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA.s ~ Shortfall.mn.2 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| 0 |0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s ~ Shortfall.mn.2 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| Country |0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.total ~ Shortfall.mn.2 
                 | 0 |0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.RWA.hat ~ Shortfall.mn.2 
                 | 0 |0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(out.CAR.ratio ~ Shortfall.mn.2 
                 | 0 |0|0,
                  data = DiD.data)
se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(out.tier.ratio ~ Shortfall.mn.2 
                 | 0 |0|0,
                  data = DiD.data)
se.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```

```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}

stargazer(reg.1, reg.4, reg.5, reg.6, reg.7,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is the capital shortfall during the capital exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA average", "$\\Delta$ Capital ratio", "$\\Delta$ CET1 ratio"),
          se = list(se.1, se.4, se.5, se.6, se.7),
          covariate.labels = c("Log Capital Shortfall"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Country FE", "\\multicolumn{1}{c}{No}", "\\multicolumn{1}{c}{No}",    
                             "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))

```   
```{r, include=FALSE}

reg.1 <- felm(out.RWA.s ~ EBAbank | 0 |0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA.s ~ EBAbank 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| 0 |0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s ~ EBAbank 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| Country |0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.total ~ EBAbank 
                 | 0 |0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.RWA.hat ~ EBAbank 
                 | 0 |0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(out.CAR.ratio ~ EBAbank 
                 | 0 |0|0,
                  data = DiD.data)
se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(out.tier.ratio ~ EBAbank 
                 | 0 |0|0,
                  data = DiD.data)
se.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```

```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}

stargazer(reg.1, reg.4, reg.5, reg.6, reg.7,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings: Exercise vs Non-exercise banks",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i =\\alpha +  \\beta \\text{Capital Shortfall}_i+\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is a dummy that indicates if bank $i$ was part of the exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA average", "$\\Delta$ Capital ratio", "$\\Delta$ CET1 ratio"),
          se = list(se.1, se.4, se.5, se.6, se.7),
          covariate.labels = c("EBA"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Country FE", "\\multicolumn{1}{c}{No}", "\\multicolumn{1}{c}{No}",    
                             "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))

```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier,log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Ratio Savings and Bank Characteristics",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Capital ratio savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{CT1 ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. $\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```
```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif  | year +  IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Ratio Savings and Bank Characteristics: Bank $\\times$ IRB Fixed effects",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times\\text{Capital ratio savings}_{it}+\\alpha_{i}\\times\\text{IRB}_{it}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. $\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ CAR.dif + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Ratio Savings and Bank Characteristics: IRB vs Non-IRB Banks",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times\\text{Capital ratio savings}_{it}+\\gamma\\text{IRB}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. $\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Capital ratio savings", "IRB","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Ratio Savings and Bank Characteristics: Only IRB Banks",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times \\text{Capital ratio savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. Only banks that use the IRB approach are included in these regressions.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "RWA Savings and Bank Characteristics",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```
```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K ,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K   ~ log.RWA.s  | year +  IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ log.RWA.s  | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ log.RWA.s  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ log.RWA.s  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ log.RWA.s  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "RWA Savings and Bank Characteristics: Bank $\\times$ IRB Fixed effects",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}\\times IRB_{it}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings. $\\text{IRB}_{it}$ is a dummy variable equal to 1 if bank i at year t used the IRB approach to calculate capital requirements and zero otherwise.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K ,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ log.RWA.s + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ log.RWA.s + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "RWA Savings and Bank Characteristics: IRB vs Non-IRB Banks",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\gamma\\times\\text{IRB}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings. $\\text{IRB}_{it}$ is a dummy variable equal to 1 if bank i at year t used the IRB approach to calculate capital requirements and zero otherwise.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Log RWA savings", "IRB","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, echo=FALSE}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(ROE  ~ log.RWA.s | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(ROA  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(Zscore  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]

reg.7 <- felm(leverage  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]

```
```{r, include=FALSE}
ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
```

```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "RWA Savings and Bank Characteristics: only IRB",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings. The sample is restricted to banks that use the IRB approach to compute their capital requirements.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```