### 2020-04-02 Analysis/Simulation/DiD-analysis.Rmd
Add backup file for the DiD analysis using the EBA capital exercise

---
title: "Credit Risk Distribution and Capital Savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn, longtable}
output:
  pdf_document:
    #citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
    fig_caption: true
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
csl: ../../Auxiliar/journal-of-finance.csl
citecolor: black
anonymous: false
fontsize: 11pt
spacing: "double"
geometry: margin= 1in
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
JEL: G21, G28, G11
linkcolor: black
biblio-style: apsr
subtitle: ''
thanks: Preliminary draft.  
abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach. Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution. Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. While a larger average component increases capital charges, a larger shape component generates capital savings. I document that these capital savings are important sources of risk-weight variability across banks. I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical. The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
urlcolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

rm(list = ls())

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions

```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

```{r did data, eval=FALSE, include=FALSE}
p_w <- c(0.0, 1)         ## Set winsorizing limits
pre.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2009","2010"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

post.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2012","2013"),
         !is.na(IRB)) %>%
  select(bvdid,  log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

DiD.data <- bank.data %>%
  filter(year %in% c("2010")) %>%
  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
         Shortfall.mn.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.mn, 0), NA),
         Shortfall.RW.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.RW, 0), NA),
         out.RWA        = log.RWA.post - log.RWA.pre,
         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
         out.RWA.s.Retail = log.RWA.s.Retail.post - log.RWA.s.Retail.pre,
         out.RWA.s.Equity = log.RWA.s.Equity.post - log.RWA.s.Equity.pre,
         out.RWA.s.Wholesale = log.RWA.s.Wholesale.post - log.RWA.s.Wholesale.pre,
         out.RWA.s.Corporate = log.RWA.s.Corporate.post - log.RWA.s.Corporate.pre,
         out.RWA.s.Sovereign = log.RWA.s.Sovereign.post - log.RWA.s.Sovereign.pre,
         out.RWA.s.Banks = log.RWA.s.Banks.post - log.RWA.s.Banks.pre,
         out.K          = log.K.post - log.K.pre,
         out.tier       = log.tier.post - log.tier.pre,
         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
         out.tier.dif   = tier.dif.post - tier.dif.pre,
         Country        = as.factor(Country)) %>% 
  ungroup() %>% 
  filter(EBAcountry == 1,
         name != "Bank of Cyprus") %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

```

```{r did, eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
p_w <- c(0.0, 1)         ## Set winsorizing limits
DiD.data %>% ungroup()%>%
  subset(select = c("Shortfall.mn","out.RWA.s","out.RWA.s.Retail","out.RWA.s.Wholesale","out.RWA.s.Corporate","out.RWA.s.Sovereign","out.RWA.s.Banks","out.RWA", "out.K","out.CAR.ratio","out.CAR.dif")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Log Shortfall","$\\Delta \\text{Log RWA}^S$", "$\\Delta \\text{Log RWA}^S_R$", "$\\Delta \\text{Log RWA}^S_W$", "$\\Delta \\text{Log RWA}^S_C$", "$\\Delta \\text{Log RWA}^S_S$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}$", "$\\Delta \\text{Log Capital}$", "$\\Delta \\text{CAR}^S$"  ),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

```{r, eval = FALSE, fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
 plot1 <- bank.data %>%  ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier, tier.hat, IRB) %>%
  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, CAR, CAR.hat, IRB) %>%
  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, CAR, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

```{r,eval=FALSE, echo=FALSE}

bank.data.2 <- bank.data %>% ungroup()%>%
 # filter(!is.na(RW.s),
  #       !is.na(basel),
   #      tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  RW, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
#  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif  | year + IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(log.asset  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(RW  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Savings and Capital Ratio Components",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times \\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5),
          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","Log Assets","RW"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r,eval=FALSE, echo=FALSE}

bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Savings, Performance, and Risk-Taking",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times\\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. $\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4),
          covariate.labels = c("Capital ratio savings"),
          dep.var.labels = c("ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, eval=FALSE, echo=FALSE}

reg.1 <- felm(out.tier.ratio ~ Shortfall.mn
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA ~ Shortfall.mn 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s ~ Shortfall.mn 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.hat ~ Shortfall.mn 
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.tier ~ Shortfall.mn
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is the capital shortfall during the capital exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$\\Delta$ CET1 ratio", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA Counterfactual", "$\\Delta$ Log CET1" ),
          se = list(se.1, se.2, se.3, se.4, se.5),
          covariate.labels = c("Log Capital Shortfall"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE)

```

```{r,eval=FALSE, echo=FALSE}

reg.1 <- felm(out.RWA.s ~ Shortfall.mn 
               |0|0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA.s.Retail ~ Shortfall.mn 
                |0|0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s.Wholesale ~ Shortfall.mn 
                |0|0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.s.Corporate ~ Shortfall.mn
                |0|0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.RWA.s.Sovereign ~ Shortfall.mn
                |0|0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(out.RWA.s.Banks ~ Shortfall.mn 
               |0|0|0,
                  data = DiD.data)
se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
```

```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings: Portfolio Breakdown",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^s_{i,j} = \\alpha + \\beta \\text{Log Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is the capital shortfall during the capital exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$j = \\text{Total}$", "$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"),
          #se = list(se.1, se.2, se.3, se.4, se.5, se.6),
          covariate.labels = c("Log Shortfall"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
         # float.env = "sidewaystable",
          align = TRUE)
```