### 2020-07-17 Analysis/Draft/Draft_bookdown.Rmd
Draft version of first revision by supervisors

---
title: "The Countercyclical Component of Capital Regulation"
date: '\emph{\today}'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    #citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
  pdf_document: default
  word_document: default
author:
- name: "Lucas Avezum"
  affiliation: "CentER and EBC, Tilburg University"
biblio-style: apsr
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
csl: ../../Auxiliar/Citation_styles/american-marketing-association.csl
#anonymous: no
fontsize: 11pt
geometry: margin= 1in
header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float, siunitx, tabularx, appendix} 
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
linkcolor: black
abstract: The Basel II accord introduced a concave mapping from probabilities of default
  (PD) to capital requirements for banks adopting the internal ratings-based (IRB)
  approach. Therefore, total capital requirements depends not only on the portfolio's
  average PD but also on the shape of its PD distribution. Using a hand-collected
  dataset, I propose a method that decomposes the contribution of the PD distribution
  to capital requirements into an average component and a shape component. While a
  larger average component increases capital charges, a larger shape component generates
  capital savings. I investigate the relationship of these capital savings with the business cycle and find it to   be countercyclical. The results suggest that a recalibration of the current IRB framework can reduce considerably the procyclicality of capital regulation.
spacing: double
indent: true
subtitle: ''
thanks: Preliminary draft.
JEL: G21, G28, G11
urlcolor: black
toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE,
                      fig.pos   = 'H')
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

rm(list = ls())

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(multipanelfigure)     ## Create grid for figures
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
library(splitstackshape)      ## Command to expand data
library(latex2exp)            ## Use latex code in plots
library(lemon)                ## Plot brackets
library(mosaic)               ## Package to calculate derivatives
#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions
source("../../Auxiliar/longtable.R") ## longtable for stargazer
#source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer

```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

\section{Introduction} \label{sec:intro}

The regulation on bank capital requirements has evolved substantially since the first international standard, the 1988 Basel Accord. At the heart of these changes is the link of capital charges to asset risk [@BCBS2006international; @BCBS2017basel; @EBA2019policy]. Risk sensitivity of capital requirements was present in the first accord and it was greatly enhanced in the Basel II framework by giving banks the option to calculate capital requirements using risk estimates obtained from their own models, known as the internal ratings-based (IRB) approach. The rationale for risk-linked capital charges is that regulating capital via a simple capital to asset ratio can incentivize banks to take on excessive risk [@koehn1980regulation; @kim1988risk; @gordy2010risk]. The rationale for allowing internal models is the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].

However, both risk sensitivity and the use of internal models may have adverse consequences. First, linking capital charges to asset risk exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical]. As the overall credit quality deteriorates during economic downturns, capital requirements tighten, forcing banks to reduce lending. Second, the inherent flexibility of the IRB approach may allow for differences in capital requirements across banks that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits;@berg2017analysis; @ferribank].

This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements. Under the IRB approach, banks supply their own risk estimates, such as the probability of default (PD), to formulas set by the regulation which determine the amount of capital banks have to hold. Two characteristics of the framework are important for this analysis. First, the mapping from PD estimates to capital requirements is a concave function. Second, capital requirements are calculated for each asset separately and then added up to obtain the total capital charge. Because of these characteristics, actual capital requirements for a portfolio are lower than implied by the average PD of this portfolio. For clarity, consider two portfolios. The first contains a low PD asset and a high PD asset in equal proportions. The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD. Because of the concavity of the IRB formula, capital requirements for the second portfolio are greater than capital requirements for the first. More generally, total capital requirements depend not only on banks' portfolio average PD but also on the shape of the PD distribution.

I propose a measure that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. The average component is a counterfactual scenario such that the capital requirements for the entire portfolio are calculated using the portfolio’s average PD. The shape component is obtained as the difference between the average component and the actual amount of capital requirements. I label the shape component *Capital savings* because an increase of the shape component reduces total capital requirements.

To calculate *Capital savings*, I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports. The sample consists of 59 large banking groups that have adopted the IRB approach. The dataset covers 15 countries and goes from 2007 to 2018.

With this novel dataset, I investigate the relationship between *Capital savings* and the business cycle. Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD estimates is monotonic. Conversely, the direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle. The ambiguity arises because *Capital savings* can only increase during a recession, and therefore be countercyclical, if the shape of the PD distribution is affected during the business cycle. If the PD estimates of the the assets in a portfolio are equally affected during the business cycle, then there is only a shift of the location parameter of the PD distribution. Because the PD elasticity of capital requirements is lower for higher PD estimates, in this scenario *Capital savings* are procyclical. On the other hand, if riskier assets are sufficiently more sensitive to credit shocks than safer assets, then *Capital savings* are countercyclical.

I find *Capital savings* to be countercyclical. I estimate that a 1.7% GDP growth rate (one standard deviation) decreases the growth rate of *Capital savings* by 0.8 percentage points on average. In other words, given an average PD, banks have to increase capital by 0.8 percentage points to keep their capital ratio constant. The effect is also economically significant. The variation of *Capital savings* during the business cycle reduces the procyclicality of capital ratio by 23.7% on average. The results also suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios. The results are robust to different measures of the shape component, to the inclusion of several control variables and the alternative sample selections. I also show that selection into the IRB approach does not explain the cyclicality of *Capital savings*.

I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic recessions average and variance of PD distributions increase while the their skewness decrease. Although the effect of the average PD is stronger, which characterizes capital requirements as procyclical, both the higher variance and the lower skewness of the PD distribution mitigate the increase of capital requirements during recessions. Moreover, I argue that the change of skewness during the business cycle is consistent with a portfolio reallocation effect on *Capital savings*.

This papers contributes to two strands of the banking literature. First, several studies document significant capital requirements variability across otherwise similar banks. Some take this evidence as indicative of either excessive subjectivity in the current capital requirement rules or risk measurement manipulation by banks to decrease capital charges [see, @mariathasan2014manipulation; @ferribank]. Others are more cautious and emphasize that part of the capital requirements variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory;@berg2017analysis; @EBA2019result]. I contribute to the literature by showing that *Capital savings* are an important source of capital requirements variation across banks.

Second, while the literature shows that capital requirements, especially for portfolios under the IRB approach, are on average procyclical [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical], I identify a countercyclical component of the current IRB framework. A better understanding of this countercyclical component can help policy makers designing a less procyclical regulation in the future.

The remainder of this paper is organized as follows. Section \ref{sec:background} describes the capital requirement formula under the IRB approach, why it creates *Capital savings*, and the hypothesis concerning the cyclicality of *Capital savings*. Section \ref{sec:data} describes the collected data and the measures of *Capital savings*. Section \ref{sec:cycle} presents evidence on the countercyclical feature of *Capital savings*. Section \ref{sec:reallocation} tests if *Capital savings* are affected by portfolio reallocation during the business cycle. Section \ref{sec:additional_results} presents additional results. Finally, section \ref{sec:conclusion} concludes.

\section{Institutional Background and Hypotheses} \label{sec:background}

\subsection{Capital Requirements under the IRB approach} \label{sec:cap_req}

In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@BCBS2006international], known as Basel II, which serves as the basis for national rulemaking and implementation processes. The link of capital charges to asset risk was at the heart of the revision. Since the first accord from 1988 (Basel I), minimum capital requirements are calculated as a percentage of risk-weighted assets (RWAs). Hence, choosing the approach that defines risk-weights is crucial. In Basel I, RWs were constant within portfolio categories. For instance, all sovereign exposures were weighted by 0% and all corporate exposures by 100%. However, regulating capital via a simple capital to asset ratio can incentivize banks to take on excessive risk as the cost in terms of capital requirements is the same within a portfolio category [@koehn1980regulation; @kim1988risk; @gordy2010risk]. Therefore, since Basel II, financial institutions may choose between two approaches to calculate capital requirements for credit risk; the standardized approach (essentially a slightly modified version of the first accord) and the IRB approach.

Under the IRB approach, banks are required to estimate their credit risk parameters using internal risk models. The risk parameters are the probability of default (PD), the loss given default (LGD), and the exposure at default (EAD). The main rationale for allowing internal models is the better alignment of regulatory risk estimates to actual portfolio risk [@barakova2014banks] as banks are expected to have better knowledge of and more resources to monitor their portfolio than supervisors have. Because the internal models have to be approved by the supervisors before implemented for regulatory purposes, regulators also expected the IRB approach to promote stronger risk management practices in the banking industry [@BCBS2006international; @cucinelli2018credit].

The estimated risk parameters are then used in regulatory formulas for calculating risk-weights. The regulatory formulas comes from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e. the model is portfolio invariant. This feature leads to a bottom-up approach where capital requirements are determined on the asset level and total requirement is simply the sum of asset individual requirement. Equation \ref{eq:RW_function} is the risk-weight formula at its most simple form:

\begin{equation}
\label{eq:RW_function}
\text{RW}(\text{LGD}_i, \text{PD}_i)= 12.5\text{LGD}_i\bigg[ N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\text{PD}_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)-\text{PD}_i\bigg],
\end{equation}
where the subscript $i$ indexes an asset in the portfolio, $N$ is the standard normal cumulative distribution, and $R$ is a correlation coefficient. Basel accords segment credit portfolios into several categories each having different additional features while keeping the basic structure of equation \ref{eq:RW_function}. For instance, the correlation coefficient is a function of PD for wholesale exposures while it is constant at 15% for real estate exposures.[^1] Although the regulatory formulas are common for all banks the scope that banks are allowed to use their internal models varies. Under the foundation IRB (F-IRB) approach, banks can apply or be approved to use only internal PD estimates and are required to use values defined by the regulator for the other parameters. Only banks approved to the advanced IRB (A-IRB) approach can use internal estimates for all the risk parameters.

Because of the portfolio invariant feature of the model, average risk-weight, $\text{RW}^f$, is the exposure weighted average of all individual assets' risk-weights, as shown in equation \ref{eq:RW_factual}:[^2]

\begin{equation}
\label{eq:RW_factual}
\text{RW}^f= \sum_{i}q_i \text{RW}(\text{LGD}_i, \text{PD}_i).
\end{equation}
where $q_i= \text{EAD}_i/\sum_j\text{EAD}_j$. The total amount of RWAs for this portfolio is $\text{RW}^f\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.

[^1]: Some other differences are maturity adjustments for wholesale exposures and the use of annual sales turnover to adjust the correlation coefficient for SMEs.

[^2]: The superscript $f$ (for factual) is used to differentiate the average value from the RW function.

\subsection{Probability of Default Distribution and Capital Savings} \label{sec:PD_capital_savings}

Figure \ref{fig:pdmap} plots the mapping from PD to risk-weights according to equation \ref{eq:RW_function}. For the purpose of this paper, the important characteristic of this mapping is that it is a concave function. The concavity of the regulatory formula together with the bottom-up method to aggregate capital requirements entail that total capital requirements for a portfolio do not depend only on the portfolio average PD but also on the shape of the PD distribution. For clarity, consider two portfolios. The first contains a low PD asset and a high PD asset in proportions $q$ and $1-q$. The second has the same amount of exposure as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD weighted by their respective shares. Because of the concavity of the IRB formula and Jensen’s inequality, the capital requirement for the second portfolio is greater than the first. Figure \ref{fig:pdmap} illustrates the comparison. $\text{RW}^f$ and $\text{RW}^c$ are the average risk-weight for the first portfolio and the second portfolio, respectively. More generally, any PD distribution with positive variance is charged with capital requirements lower than implied by the average of the distribution.   
I propose a method that decomposes the contribution of the PD distribution to capital requirements into the average component and the shape component. The method is a generalization of the example in Figure \ref{fig:pdmap}. The average component is the amount of capital requirements if the portfolio’s average PD is used for the entire portfolio. Similar to $\text{RW}^c$ in Figure \ref{fig:pdmap}, the average component is defined as the following:

\begin{equation}
\label{eq:RW_counterfactual}
\text{RW}^c= \sum_{i}q_i \text{RW}(\text{LGD}_i,\overline{\text{PD}}),
\end{equation}
where the PD parameter for all assets is replaced by the portfolio weighted average. The latter is defined as the following:

\begin{equation*}
\overline{\text{PD}}=\frac{\sum_i\text{EAD}_i\text{LGD}_i\text{PD}_i}{\sum_i\text{EAD}_i\text{LGD}_i}.
\end{equation*}

The difference between $\text{RW}^c$ and $\text{RW}^f$ measures any contribution to capital requirements from the PD distribution that is not related to the location of this distribution, i.e., it is a measure of the shape component. It can also be interpreted as the amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula and the shape of the PD distribution. I define this intensity measure of *Capital savings* as the following:

\begin{equation}
\label{eq:RW_savings}
\text{RW}^{s}=\text{RW}^c-\text{RW}^f.
\end{equation}

The total amount of *Capital savings* is simply $\text{RW}^{s}$ multiplied by total EAD in the portfolio as follows:

\begin{equation}
\label{eq:measure2}
\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}^f=\text{RW}^{s}\times \sum_i\text{EAD}_i.
\end{equation}

Because this is a measure in monetary units, $\text{RWA}^{\$}$ is mechanically larger for larger banks and they increase as the portfolio rolls out to the IRB approach.

An alternative way to measure the intensity of *Capital savings* is the ratio of equation \ref{eq:RW_counterfactual} to equation \ref{eq:RW_factual} as follows:

\begin{equation}
\label{eq:measure1}
\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}^f}.
\end{equation}

$\text{RWA}^s$ is the rate at which a bank saves capital. For instance, a value of 1.5 means this portfolio would require 50% more capital in the absence of *Capital savings* to keep the same capital ratio. A value of one means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution. Any PD distribution with positive variance results in positive *Capital savings*.

$\text{RWA}^s$ is also my preferred measure of *Capital savings* because it is the most straightforward measure to evaluate the impact of *Capital savings* on the capital ratio during the business cycle. In the empirical analysis, I study the behavior of variables during the cycle by relating their growth rates to GDP growth rate. Hence, I regress the change in the logarithm of my variables of interest on GDP growth rate. Taking my empirical approach into account, $\text{RWA}^s$ becomes the natural measure of *Capital savings* as the following decomposition of the log of capital ratio shows:

\begin{equation}
\begin{aligned}
\label{eq:decomposition}
\text{Log}\bigg(\frac{\text{Capital}}{\text{RWA}^f}\bigg) &= \text{Log Capital} -  \text{Log RWA}^f \\
&= \text{Log Capital} -  \text{Log}\bigg(\frac{\text{RWA}^c}{\text{RWA}^s}\bigg) \\
&= \text{Log Capital} -  \text{Log RWA}^c + \text{Log RWA}^s
\end{aligned}
\end{equation}

If I regress all the change of these four variables - $\frac{\text{Capital}}{\text{RWA}^f}$, $\text{Capital}$, $\text{Log RWA}^c$, and $\text{Log RWA}^s$ - on GDP growth rate I obtain four estimates $\hat\beta_{\text{Cap. ratio}}$, $\hat\beta_{\text{Capital}}$, $\hat\beta_{\text{RWA}^c}$, and $\hat\beta_{\text{RWA}^s}$, respectively. Using the predicted values implied by these estimated in Equation \ref{eq:decomposition} we get the following  equality: $\hat\beta_{\text{Cap. ratio}} = \hat\beta_{\text{Capital}} -\hat\beta_{\text{RWA}^c} + \hat\beta_{\text{RWA}^s}$, which can be used to obtain the relative importance of changes of each of these three variables to changes of the capital ratio during the business cycle.

Note that I defined *Capital savings* such that positive values translate to lower actual capital requirements and consequently $\text{RW}^s \geq 0$, $\text{RWA}^{\$} \geq 0$, and $\text{RWA}^s \geq 1$.

Lastly, for robustness of the empirical analysis, I also calculate the Gini coefficient of the PD distribution as an alternative measure to my approach.

\subsection{Hypotheses on Capital Savings during the Business Cycle} \label{sec:hypotheses}

Overall, capital requirements under the IRB approach are procyclical. During economic downturns credit risk increases which raises capital requirements forcing banks to either raise equity or cut down lending. Theory and empirical evidence suggest that banks opt for the latter as raising equity is costly, even more so during recessions [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].

Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD is negative and monotonic. Consider again the example shown in Figure \ref{fig:pdmap} with a portfolio containing a share $q$ of a safe asset with PD equal to $\text{PD}_L$ and a share $1-q$ of a risky asset with PD equal to $\text{PD}_H$ such that $\text{PD}_H>\text{PD}_L$. For practical reasons I restrict the analysis to the  $\text{PD} \in [0,0.25)$. Figure \ref{fig:densities} in Appendix \ref{sec:appendixa} shows that the percentage of PD within the $[0,0.25)$ interval is between 99.35% and 99.98%. Figure \ref{fig:derivatives} in Appendix \ref{sec:appendixa} shows that for $\text{PD} \in [0,0.25)$, $\frac{\partial\text{RW}}{\partial\text{PD}}>0$, $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}<0$, and $\frac{\partial^3\text{RW}}{\partial\text{PD}^3}>0$. I assume, without loss of generality, a constant $\text{LGD}_i=\text{LGD}$ and constant $q$. Using the weighted average of this portfolio in equation \ref{eq:RW_counterfactual}, the average component for this portfolio is equal to the following:

\begin{equation}
\label{eq:RW_counter_example}
\text{RW}^{c}(\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H).
\end{equation}

Because $\frac{\partial\text{RW}}{\partial\text{PD}}>0$ and as long PD estimates are affected monotonic and negatively by GDP growth, the average component of capital requirements increases during economic downturns.

Conversely, the direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle.  The ambiguity arises because *Capital savings* can only increase during a recession, and therefore be countercyclical, if the shape of the PD distribution is affected during the business cycle. If the PD estimates are equally affected during the business cycle, then there is only a shift of the location parameter of the PD distribution. Because the PD elasticity of capital requirements is lower for higher PD estimates, in this scenario *Capital savings* are procyclical. On the other hand, if riskier assets are sufficiently more sensitive to credit shocks than safer assets, then *Capital savings* are countercyclical. To illustrate the ambiguity, I consider three cases.

First, note that using the example in Figure \ref{fig:pdmap}, *Capital savings* can be written as the following:

\begin{equation}
\label{eq:RW_savings_example}
\text{RW}^{s}(\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}(\text{PD}_L)+(1-q)\text{RW}(\text{PD}_H).
\end{equation}

We can find the effect of changes in the PD distribution to *Capital savings* by taking the total derivative of \ref{eq:RW_savings_example}:

\begin{equation}
\label{eq:RW_savings_diff}
\begin{aligned}
d\text{RW}^{s}(\text{PD}_L, \text{PD}_H)
&=\bigg[q\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}'(\text{PD}_L)\bigg]d\text{PD}_L\\
&+\bigg[(1-q)\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-(1-q)\text{RW}'(\text{PD}_H)\bigg]d\text{PD}_H\\
\end{aligned}
\end{equation}

Next, consider the case that all assets are affected equally during the business cycles in term of PD, i.e.,  $d\text{PD}_H=d\text{PD}_L=d\text{PD}$. In this scenario equation \ref{eq:RW_savings_diff} becomes:

\begin{equation}
\label{eq:first_case}
\begin{aligned}
d\text{RW}^{s}
&=\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\bigg(q\text{RW}'(\text{PD}_L)+(1-q)\text{RW}'(\text{PD}_H)\bigg)\bigg]d\text{PD}.\\
\end{aligned}
\end{equation}

Because $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}$ is a convex function and Jensen's inequality, if $d\text{PD}_H=d\text{PD}_H=d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}<0$. In this case, *Capital savings* are procyclical, decreasing during economic downturns. Figure \ref{fig:pdmapcasesymm} shows graphically this first case.

Next, consider the case that riskier assets are more affected during the business cycles in term of PD, i.e.,  $\text{dPD}_H>\text{dPD}_L$. For simplicity, consider the extreme case of $\text{dPD}_H=\text{dPD}$ and $\text{dPD}_L=0$. In this scenario, equation \ref{eq:RW_savings_diff} becomes:

\begin{equation}
\label{eq:second_case}
\begin{aligned}
\text{dRW}^{s}
&=(1-q)\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_H)\bigg]d\text{PD}\\
\end{aligned}
\end{equation}

Because $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$ $\text{dRW}^{s}/\text{dPD}>0$. In this case, *Capital savings* are countercyclical, increasing when PD estimates increase. In other words, the marginal capital requirement decreases during economic downturns. Figure \ref{fig:pdmapcaseassym} shows the second case graphically.

Finally, consider the case where $\text{dPD}_H=\text{dPD}$ and $\text{dRW}^{s}=0$. Using equation \ref{eq:RW_savings_diff}, we can find $\text{dPD}_L$ satisfying these conditions:

\begin{equation}
\label{eq:third_case}
\begin{aligned}
d\text{PD}_{L}
&=d\text{PD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
\end{aligned}
\end{equation}

Because $\frac{\partial\text{RW}}{\partial\text{PD}}<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $\text{dPD}_L>0$. We also know from the first case that $\text{dPD}_L<\text{dPD}$, otherwise $\text{dRW}^{s}$ would be negative. Hence, the difference between the impact of the business cycle on riskier and safer assets have to be sufficiently large such that *Capital savings* are countercyclical. If that is not the case, then the procyclical effect on *Capital savings* shown in equation \ref{eq:first_case} dominates.

Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the shape of the PD distribution changes during the business cycle.

\section{Data and Capital Savings Measures} \label{sec:data}

To calculate the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters. However, data at such granularity is not available in the common balance sheet datasets. Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information. Therefore, to overcome data limitations, I collect information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports. The sample consists of 59 large banking groups, from 15 countries, that have adopted the IRB approach. For each bank, I collected consolidated information from the year the IRB approach was approved until 2018. The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate, sovereign, and banks for wholesale, and real estate, qualifying revolving credit, and others for retail). This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.

Figure \ref{fig:savingsbank} shows *Capital savings* calculated according to equation \ref{eq:measure1} i.e., the ratio of the average component of RWAs to actual RWAs. Each point is a bank-year observations and the bars indicates the average value across time for each bank. There is substantial variation across banks with Goldman Sachs saving the highest at an average rate of 220% and Mediobanca the lowest at an average rate of 118%. Variation across time is considerable with, for instance, *Capital savings* for Deutsche Bank going from 209% in 2010 to 161% in 2014. *Capital savings* also vary between and within countries. Figure \ref{fig:savingscountry}  aggregates the dataset by country. In the figure, each point is the average *Capital savings* for a bank and the bars are the average of across banks within a country. Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).

Figure \ref{fig:savingsportfoliobank} shows *Capital savings* for wholesale and retail portfolios using the ranking of Figure \ref{fig:savingsbank}. The Figure suggests that *Capital savings* from wholesale portfolios are positively correlated to total *Capital savings*. This correlation seems not to be true for retail portfolios.

Lastly, Figure \ref{fig:averagebank} plots, in terms of risk-weights, *Capital savings*, the average component, and total risk-weights, following equations \ref{eq:RW_savings}, \ref{eq:RW_counterfactual}, and \ref{eq:RW_factual}, respectively. Banks are ranked by total risk-weight. From equation \ref{eq:RW_savings}, the third column is obtained by subtracting the first column from the second, i.e., total risk-weights are equal to the average component minus *Capital savings*. As expected, the average component is larger than the *Capital savings* generated by the shape component. Still, because of *Capital savings*, the variation and level of total risk-weights are substantially lower than implied by the average component. Total risk-weights are 17 percentage points, on average, lower than the average component and their standard deviation is 6 percentage points lower.     
\section{Capital Savings and the Business Cycle} \label{sec:cycle}

In this section I investigate the relationship between *Capital savings* and the business cycle. I show that *Capital savings* are countercyclical, i.e., for the same average PD, capital requirements are high when the economic is booming and low during recessions.

Throughout this section, I report the results from estimating the following specification:

\begin{equation*}
\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*}

The dependent variable is the yearly change in a variable associated to portfolio $j$ for bank $i$ at year $t$. The change in the logarithm of real GDP per capita comes from the country where the bank has its headquarter. All regressions include bank and year fixed effects and standard errors are clustered at the country-level.

Table \ref{tab:main} shows the results using equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure of *Capital savings*. According to regression 1, an increase of 1.7 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.8 percentage points on average. The effect is economically sizable representing 17% of total *Capital savings* standard deviation in the sample.

The remaining columns of Table \ref{tab:main} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios. The negative estimated coefficient in Column 2 suggests that *Capital savings* generated by retail portfolios are also countercyclical. However, this coefficient is not statistically significant. On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant. According to regression 3, one standard deviation increase in GDP growth rate decreases the growth rate of *Capital savings* from wholesale portfolios by 0.9 percentage points on average. The effect is 19% of the dependent variable's standard deviation. Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios. Only in regression 4 the coefficient of GDP growth rate is both negative and statistically significant.

Next, I evaluate the impact of *Capital savings* on the capital ratio during the business cycle. I regress the yearly change of each capital ratio component on GDP growth rate. As explained in Section \ref{sec:PD_capital_savings}, the effect of GDP growth rate on each component sum up to the effect on the capital ratio. Consequently, the contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.

Table \ref{tab:contribution} shows that the growth rate of *Capital savings* is the only countercyclical component of capital ratio growth rate. As expected, both the amount of capital and RWAs are procyclical which translate to capital ratios also being procyclical. The results also confirm the expectation that the procyclical effect of the average component dominates over the countercyclical *Capital savings*. Nevertheless, the contribution of *Capital savings* is economically relevant, reducing the procyclicality of capital ratios by 23.7\% on average.[^3]

[^3]: The coefficient on total *Capital savings* differs slightly from Table \ref{tab:main} because in Table \ref{tab:contribution} the sample is restrict to banks with observable capital ratio and is winsorized at 1% instead of 5%. The process of winsorizing is also the reason why the contributions do not add-up to 100%.

\subsection{Robustness to Alternative Measures and Samples} \label{sec:robustness}

In this section, I test the robustness of the main results to alternative measures of *Capital savings*. In Panel A of Table \ref{tab:alt}, the dependent variable measures the amount *Capital savings* in US dollars following equation \ref{eq:measure2}. As before, the estimated coefficients suggest that the countercyclical effect is mostly driven by wholesale portfolios. Differently from Table \ref{tab:main}, the coefficient of GDP growth rate estimated on the bank portfolio is now negative and statistically significant. Because coefficients in Panel A of Table \ref{tab:alt} are  impacted by the size of the portfolio, while in \ref{tab:main} they are not, the results of regressions 6 in both tables suggest that part of the *Capital savings* countercyclicality comes from an increase in exposure to other banks during recessions.

In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:RW_savings}. This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity — i.e., risk-weights. Although the coefficient on the wholesale portfolio is no longer statistically significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1% confidence level. The attenuation of the effect on wholesale portfolio appears to come from large, albeit not statistically significant, procyclical *Capital savings* in sovereign portfolios. Moreover, the small and statistically insignificant coefficient in regression 6 reinforces the claim that countercyclical *Capital savings* in the bank portfolio are mostly due to the extensive and not the intensive margin.

In Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution. Again, the results in Panel C support the evidence from Table \ref{tab:main} i.e., *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.

Lastly, Appendix \ref{sec:appendixc} shows that the results are robust to the addition of bank- and country-level controls (Table \ref{tab:extension}), to including only banks with at least 10 observations (Panel A of Table \ref{tab:robustness}), and to winsorizing the sample to 1% level instead of 5% (Panel B of Table \ref{tab:robustness}).

\section{Capital Savings and Portfolio Reallocation} \label{sec:reallocation}

In this section I argue that portfolio reallocation during the business cycle explains part of the *Capital savings* cyclicality. As method, I study the changes of the PD distribution moments during the business cycle and check if these changes are consistent with a portfolio reallocation effect on *Capital savings*.[^4]

[^4]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PD estimates evolution to the PD estimates of new loans. Unfortunately, in my dataset, I only observe assets aggregated by PD bands without knowledge if amount changes in a band are due to new loans or changes in credit quality of assets already held.

According to the theoretical framework presented in section \ref{sec:hypotheses}, an increase in average PD reduces the amount of *Capital savings* holding other moments constant. This relationship is illustrated in Figure \ref{fig:pdmapcasesymm} and empirically tested in Table \ref{tab:moments_contribution}. In all columns of this table average PD is negatively associated with *Capital savings*. Note that, in Table \ref{tab:moments_contribution}, not only the explained but also explanatory variables change across regressions by portfolio $j$.

Panel A in Table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle. As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth. This association is the driver of the procyclical character of the current model-based capital regulation. The only exception is the sovereign portfolio. Column 5 of Panel A shows that the coefficient of GDP growth rate on changes in the average PD of sovereign portfolios is positive, although not statistically different from zero. A possible explanation for the positive and insignificant result is that during economic downturns banks may search for safe assets mostly in the form of safe haven government bonds. According to column 5 of Panel A, this reallocation of portfolio appears to be sufficient to compensate any deterioration of the prior held portfolio. Together with column 5 in Table \ref{tab:moments_contribution}, Column 5 of Panel A in Table \ref{tab:moments_cyclicality} also helps explain the positive and insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.

Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios. And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, Panel B in Table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios. In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and Panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section \ref{sec:hypotheses}.

At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted exactly because the estimated effect of these two moments on *Capital savings* are strongly aligned with the theoretical framework, which does not allow for portfolio reallocation. Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness. It is trivial that the equivalent case with a continuous PD distribution to the scenario in Figure \ref{fig:pdmapcasesymm} would not affect the skewness of the distribution. Still in this case, only a change in the location parameter occurs. On the other hand, in the second scenario the PD distribution should become more skewed to the right. Even if we think that some assets will become safer during bad times, it is reasonable to expect a larger effect on the right tail than on the left tail of the PD distribution. Consequently, without portfolio reallocation, we should expect the skewness of the PD distribution to be negatively associated with the GDP growth rate.

Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle is consistent with portfolio reallocation. All the statistical significant coefficients are positive, contrary to the expectation if banks would not reallocate their portfolio. The positive coefficients suggest that banks respond to the deterioration of their portfolios' credit quality by reducing risk taking in new loans. The combination of these two movements, higher PD of existing exposures and lower PD of new assets, results in a less skewed PD distribution during recessions.

Table \ref{tab:moments_contribution} shows that lower skewness translates into higher *Capital savings*. The effect is statistical significant for all except sovereign portfolios. According to column 1, an increase of 18.2 percentage points in PD distribution skewness growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.9 percentage points on average, representing 20% of total *Capital savings* standard deviation. The effect is even stronger on corporate portfolios accounting for 41% of their *Capital savings* standard deviation.

The evidence of Panel C in Table \ref{tab:moments_cyclicality} and Table \ref{tab:moments_contribution} suggest that part of the increase in *Capital savings* in economic recessions is consistent with a portfolio reallocation effect. This countercyclical effect, driven by the skewness of the PD distribution, is on top of any gain from a lower average PD also a result of the reduced risk taking.

\subsection{Selection and Reallocation Across Portfolios} \label{sec:cross_reallocation}

In this section, I show that selection into IRB is not driving my results. Although banks cannot opt out of IRB approach once it has been approved, they can try to (re)allocate assets to portfolios under the less risk-sensitive standardized approach. Consequently, the countercyclical effect of *Capital savings* can potentially be confounded with this across approach reallocation.

Table \ref{tab:other} provides little support for cross approach reallocation during the business cycle. The growth rate of the total amount (column 1), the average risk-weight (column 2), and the share under the standardized approach (column 3) are all positively associated with GDP growth rate but none of the coefficients are statistically significant. The positive coefficient in column 3 is also not in line with the expectation that banks would move assets to the standardized approach during economic downturns. The positive coefficient in column 1 is most likely driven by the balance sheet expansion during economics booms while the positive coefficient in column 2 is reflecting portfolio reallocation within the standardized approach during the business cycle, for instance, from sovereign to corporate lending.

Lastly, I test if banks reallocate across portfolios under the IRB approach during the business cycle. According to columns 4 and 5 in Table \ref{tab:other}, it seems that banks increase retail exposure at the expense of wholesale exposure during periods of economic expansion. However, I find this reallocation to be statistically insignificant.

Overall, Table \ref{tab:other} supports ruling out selection concerns to the result that *Capital savings* are a countercyclical component of capital requirements.

\section{Conclusion and Discussion} \label{sec:conclusion}

This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements. *Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are countercyclical and economically relevant. *Capital savings* reduce the procyclicality of the capital ratio by 23.7%. Overall, this paper provides evidence that the concavity of IRB formulas could be explored to reduce the procyclicality of current capital regulation.

\section*{References} 
\noindent
<div id="refs"></div>

\newpage

# Figures {-}

```{r parameters curve, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 1
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
data <- mapping_retail(grid,1,0.15,0.999)
PD.l <- 0.01
PD.h <- 0.15
PD.a <- (PD.l+PD.h)/2
PD.h2 <- 0.17
PD.h3 <- 0.16
PD.l3 <- 0.02
PD.a2 <- (PD.l+PD.h2)/2
PD.i  <- 0.19
K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
K.i  <- as.numeric(mapping_retail(PD.i,1,0.15,0.999)["K"])
K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
```

```{r, results="hide"}
pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from PD to risk-weights according to equation \\ref{eq:RW_function}. $LGD$ is set equal to one and the correlation coefficient to 0.15."

```

```{r, fig.height = 3, fig.width = 4,results="hide", fig.cap = pd_map_cap, fig.align = 'center', echo=FALSE, message=FALSE}
p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+

  theme_bw()+
  theme(#axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_text( hjust = 1),
        axis.title.y = element_text( hjust = 1, angle = 0),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous(name= "PD", 
                     breaks=c(PD.l,PD.a,PD.h),
                     labels= unname(TeX(c("$PD_{L}", "$\\bar{PD}", "$PD_{H}"))),
                     expand = c(0, 0))+
  scale_y_continuous(name= "RW",
                     breaks=c(K.s,K.a),
                     labels=unname(TeX(c("$RW^{f}", "$RW^{c}" ))),
                     expand = c(0, 0)) 
```

```{r parameters curve 2, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 1
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
data <- mapping_retail(grid,1,0.15,0.999)
PD.l <- 0.01
PD.h <- 0.13
PD.a <- (PD.l+PD.h)/2
PD.h2 <- 0.19
PD.h3 <- 0.16
PD.l3 <- 0.04
PD.a2 <- (PD.l+PD.h2)/2
PD.l.a <- (PD.l+PD.l3)/2
PD.h.a <- (PD.h+PD.h3)/2
PD.h.a2 <- (PD.h+PD.h2)/2
K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
```

```{r}
cyclicality_cap_symm = "\\label{fig:pdmapcasesymm}The figure shows how changes in PD can lead to procyclical \\textit{Capital savings}. The dotted lines correspond to the portfolio before the credit risk shock and the dashed lines to the portfolio after the the shock. The arrow segments in the plot show the size of the capital savings before the shock."
```

```{r, cyclicalitysymm, fig.height = 3, fig.width = 4,results="hide", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap_symm}

 ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K))+ 
  geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+
  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a, y=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a+0.081, yend=K.s+0.081),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a+0.081, y=K.s+0.081),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+

  theme_bw()+
  theme(#axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_text( hjust = 1),
        axis.title.y = element_text( hjust = 1, angle = 0),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
 scale_x_continuous(name = "PD", breaks=c(PD.l.a,PD.h.a),labels=c("dPD","dPD"), expand = c(0, 0))+
  scale_y_continuous(name = "RW", breaks = NULL, expand = c(0, 0)) +
coord_flex_cart(bottom = brackets_horizontal(length = unit(0.075, "npc")))
```
```{r}
cyclicality_cap_asym = "\\label{fig:pdmapcaseassym}The figure shows how changes in PD can lead to countercyclical \\textit{Capital savings}. The dotted lines correspond to the portfolio before the credit risk shock and the dashed lines to the portfolio after the the shock. The arrow segments in the plot show the size of the capital savings before the shock. "
```

```{r, cyclicalityasym, fig.height = 3, fig.width = 4,results="hide", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap_asym}

ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K))+ 
  geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+
  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a, y=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l+0.01, xend=PD.l+0.01, y=K.a+0.025, yend=K.s+0.025),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
  geom_segment(aes(x=PD.l+0.01, xend=PD.l+0.01, yend=K.a+0.025, y=K.s+0.025),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
 theme_bw()+
  theme(#axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_text( hjust = 1),
        axis.title.y = element_text( hjust = 1, angle = 0),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
 scale_x_continuous(name = "PD",breaks=c(PD.h.a2),labels=c("dPD"), expand = c(0, 0))+
  scale_y_continuous(name = "RW",breaks=NULL,expand = c(0, 0)) +
  coord_flex_cart(bottom = brackets_horizontal(length = unit(0.15, "npc")))

```

```{r}
savings_cap = "\\label{fig:savingsbank}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure. Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
```

```{r, include = TRUE, fig.height = 7.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r.Total)%>%
  group_by(name, Country) %>%
#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
  group_by(Country) %>%
 # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
  #ungroup()%>%
  #mutate(ordering = order_aux+RWA.r,
  #       name = as.factor(name),
  #       name = fct_reorder(name, ordering),
  #       Country = fct_reorder(Country, ordering)) %>%
 ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
                      y = RWA.r.Total)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
    coord_flip(ylim=c(1, 2.5)) +
  labs(y =  unname(TeX(c("$RWA^{s}$" )))) +
  scale_fill_grey()+
    theme_bw()

```

```{r}
savings_cap_country = "\\label{fig:savingscountry}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure. Each point is the across time average \\textit{Capital savings} for a bank and the bars are the average across banks within a country."
```

```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
  select(country.name,name, year, RWA.r.Total)%>%
  group_by(country.name,name) %>%
 summarise_if(is.numeric,mean,na.rm= TRUE)%>%
 ggplot(mapping = aes(x = reorder(country.name, RWA.r.Total, FUN = mean, na.rm=TRUE),
                          y = RWA.r.Total)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
  labs(y =  unname(TeX(c("$RWA^{s}" )))) +
  coord_flip(ylim=c(1, 2.5)) +
  theme_bw()

```
```{r}
savings_portfolio_cap = "\\label{fig:savingsportfoliobank}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure. The left-hand side plots shows \\textit{Capital savings} for wholesale portfolios and the right-hand side for retail portfolios. Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
```

```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= savings_portfolio_cap, fig.align = 'center'}
bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r.Total, RWA.r.Retail, RWA.r.Wholesale)%>%
  pivot_longer(cols = c("RWA.r.Retail","RWA.r.Wholesale"), names_to = "Portfolio", values_to = "RWA.r") %>%
  mutate(Portfolio = factor(Portfolio, 
                            levels = c("RWA.r.Wholesale","RWA.r.Retail"), 
                            labels = c("Wholesale","Retail")))%>%
 ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
                      y = RWA.r)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
    coord_flip(ylim=c(1, 3.5)) +
  labs(y =  unname(TeX(c("$RWA^{s}$" )))) +
  scale_fill_grey()+
    theme_bw() +
  facet_wrap(~Portfolio)

```

```{r}
average_cap = "\\label{fig:averagebank}The figure plots, in terms of risk-weights, \\textit{Capital savings} ($\\text{RW}^s$), the average component ($\\text{RW}^c$), and total risk-weights ($\\text{RW}^f$), following equations \\ref{eq:RW_savings}, \\ref{eq:RW_counterfactual}, and \\ref{eq:RW_factual}, respectively. Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
```

```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= average_cap, fig.align = 'center'}
RW_summary <-  bank.data %>%
  mutate(RW.ranking = RW.IRB) %>%
  select(name, Country, RW.ranking, RW.s.Total, RW.hat.Total, RW.IRB)%>%
  pivot_longer(cols = c("RW.hat.Total","RW.s.Total","RW.IRB"), names_to = "Portfolio", values_to = "RW") %>%
  mutate(Portfolio = factor(Portfolio, 
                            levels = c("RW.s.Total", "RW.hat.Total", "RW.IRB"), 
                            labels = c("j = s","j = c", "j = f")))%>%
  group_by(Portfolio) %>%
  dplyr::summarize(mean = round(mean(RW, na.rm = TRUE),2),
            sd   = round(sd(RW, na.rm = TRUE),2)) %>%
  mutate(lab = paste("Mean = ", mean, "\n St. Dev. = ", sd, sep = ""))

bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
  mutate(RW.ranking = RW.IRB) %>%
  select(name, Country, RW.ranking, RW.s.Total, RW.hat.Total, RW.IRB)%>%
  pivot_longer(cols = c("RW.hat.Total","RW.s.Total","RW.IRB"), names_to = "Portfolio", values_to = "RW") %>%
  mutate(Portfolio = factor(Portfolio, 
                            levels = c("RW.s.Total", "RW.hat.Total", "RW.IRB"), 
                            labels = c("j = s","j = c", "j = f")))%>%
 ggplot(mapping = aes(x = reorder(name, RW.ranking, FUN = mean, na.rm=TRUE),
                      y = RW)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  unname(TeX(c("$RW^{j}$" )))) +
  scale_fill_grey()+
    theme_bw() +
  facet_wrap(~Portfolio)+
  geom_text(data = RW_summary, aes(label = lab), x = 3, y = 0.85, size=3.5)

```

```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
## Set winsorizing limits
p_w <- c(0.05, 0.95)        
bank.data.win5 <- bank.data %>% 
  ungroup()%>%
  mutate(year = as.character(year))%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  group_by(bvdid)%>%
  mutate(year = as.numeric(year))%>%
  mutate_at(vars(matches("log.RWA.r")), .funs = list(n = ~sum(!is.na(.))))

p_w <- c(0.01, 0.99)        
bank.data.win1 <- bank.data %>% 
  ungroup()%>%
  mutate(year = as.character(year))%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  group_by(bvdid)%>%
  mutate(year = as.numeric(year))%>%
  mutate_at(vars(matches("log.RWA.r")), .funs = list(n = ~sum(!is.na(.))))

```

\newpage

# Tables {-}

```{r tables_aux, echo=FALSE}

# Dependent variables
dep.vars <- c(".Total", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")

# Independent variables
ind.vars <- c("log.gdp_gr")
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")

fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

```

```{r reg_cyclicality_main, echo=FALSE}

# Model formulas
formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win5)
})

```

```{r table_cyclicality_main, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          title = "Capital Savings and the Business Cycle",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$. In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:main",
          style            = "qje",
          column.sep.width = "1pt",
          align            = TRUE)
```

```{r reg_contribution, eval=TRUE, echo=FALSE}
# Variables in the model 
# Dependent variables
dep.vars.temp <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
dep.text.temp <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{f}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")

# Model formulas
formula.text <- paste0(paste0(dep.vars.temp)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = filter(bank.data.win1, !is.na(log.CAR_gr)))
})

# Effect 
effect.list  <-  lapply(reg.list, function(x){
  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
})
effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
effect.list <-  sprintf("%.1f \\%%",effect.list)

fe.list.car <- list(c("Contribution", effect.list),
                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars.temp))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars.temp))))
```

```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.list,
          title = "Capital Savings Effect on Capital Ratio Procyclicality",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,t}^{s} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline Where $\\Delta\\text{Y}_{i,t}$ is one component of the capital ratio according to equation \\ref{eq:decomposition}. The dependent variables are the change in the logarithm of the capital ratio in column 1, of total capital in column 2, of total RWAs in column 3, of the average component in column 4, and of \\textit{Capital savings} in column 5. The average component is the amount of capital requirements if the portfolio’s average PD is used for the entire portfolio. \\textit{Capital savings} is the ratio of the average component of RWAs to actual RWAs (equation \\ref{eq:measure1}). $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. The table shows the contribution of each component to the cyclicality of capital ratios, which is the respective coefficient divided by the coefficient in column 1. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text.temp,
          covariate.labels = ind.text,
          add.lines        = fe.list.car,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:contribution",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE)

```

\begin{table}[!htbp] \centering  
\caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
\label{tab:alt}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline 
The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$. In Panel A, \textit{Capital savings} are measure in monetary units as RWAs, following equation \ref{eq:measure2}. In Panel B, \textit{Capital savings} are measure in terms of risk-weights, following equation equation \ref{eq:RW_savings}. In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}\\
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}_{i,t}$} \\
\end{tabular}

```{r reg_cyclicality_alt, echo=FALSE}

# Creating formulas and running the models
formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win5)
})

```

```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)
```

\begin{tabular}{@{\extracolsep{1pt}}cccccc}
 \\
\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}_{i,t}$ } \\
\end{tabular}
```{r reg_cyclicality_rw, echo=FALSE}

# Creating formulas and running the models
formula.text <- paste0(paste0("RW.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win5)
})

```

```{r table_cyclicality_rw, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```

\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\\
\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}_{i,t}$ } \\
\end{tabular}

```{r reg_cyclicality_gini, echo=FALSE}

# Creating formulas and running the models
formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win5)
})

```

```{r table_cyclicality_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```
\end{table}

\newpage

```{r reg_moments, echo=FALSE}

ind.text.temp <- c("$\\Delta \\text{Log mean(PD}_{i,t}^j$)", "$\\Delta \\text{Log var(PD}_{i,t}^j$)", "$\\Delta \\text{Log skew(PD}_{i,t}^j$)", "$\\Delta \\text{Kurt(PD}_{i,t}^j$)")

reg.list <- list()
# Run the models
for(i in dep.vars){
formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew | year + bvdid | 0 | Country"))
 data = rename(bank.data.win5,
               mean    = paste0("mean.PD"  , i, "_gr"),
               var     = paste0("var.PD"   , i, "_gr"),
               skew    = paste0("skew.PD"  , i, "_gr"))%>%
   felm(data = ., formula = formula.text ) -> reg.list[[i]]
  }

```

```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.list,
          title = "PD moments and Capital Savings",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log mean(PD}_{i,t}^j\\text{)}+\\gamma\\Delta \\text{Log var(PD}_{i,t}^j\\text{)}+\\kappa\\Delta \\text{Log skew(PD}_{i,t}^j\\text{)}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$. In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs. $\\Delta\\text{Log mean(PD}_{i,t}^j$) is the change in the logarithm of the average of the portfolio $j$'s PD distribution. $\\Delta\\text{Log var(PD}_{i,t}^j$) is the change in the logarithm of the variance of the portfolio $j$'s PD distribution. $\\Delta\\text{Log skew(PD}_{i,t}^j$) is the change in the logarithm of the skewness of the portfolio $j$'s PD distribution. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text.temp,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:moments_contribution",
          style            = "qje",
          column.sep.width = "1pt",
          align            = TRUE)

```

\newpage

\begin{table}[!htbp] \centering  
\caption{PD moments and the Business Cycle}  
\label{tab:moments_cyclicality}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline 
The dependent variable is the change in the logarithm of the mean in Panel A, of the variance in Panel B, and of the skewness in Panel C of the portfolio $j$'s PD distribution. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}\\
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Log mean(PD}_{i,t}^j\text{)}$} \\
\end{tabular}

```{r reg_mean, echo=FALSE}

# Model formulas
formula.text <- paste0(paste0("mean.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win5)
})

```

```{r table_mean, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```

\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\\
\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Log var(PD}_{i,t}^j\text{)}$}
\end{tabular}

```{r reg_var, echo=FALSE}

# Model formulas
formula.text <- paste0(paste0("var.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win5)
})

```

```{r table_var, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```

\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\\
\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Log skew(PD}_{i,t}^j\text{)}$}
\end{tabular}

```{r reg_skew, echo=FALSE}

# Model formulas
formula.text <- paste0(paste0("skew.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win5)
})

```

```{r table_skew, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```
\end{table}

\newpage

```{r reg_cyclicality_components, echo=FALSE}

dep.vars.temp <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
dep.text.temp <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")

# Creating formulas and running the models
formula.text <- paste0(paste0(dep.vars.temp)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

reg.list <- lapply(formulas, function(x) {
 felm(x, data = filter(bank.data.win5, year>2007))
})

```

```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.list,
          title = "Other RWA Components and the Business Cycle",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{13.5cm}{The table examines the relationship between the business cycle and other components of RWA. The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in risk-weights for portfolios under the standardize approach in column 2, the change in the ratio of exposure under the standardize approach to total exposure in column 3, the change in the wholesale share of the IRB portfolio in column 4, and the change in the retail share of the IRB portfolio in column 5. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text.temp,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "small",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:other",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE)

```

\newpage

\appendix

\section{} \label{sec:appendixa}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

```{r}
derivatives = "\\label{fig:derivatives} The figure plots risk-weights as a function of probability of default (top-left) and its first, second, and third derivative with respect to probability of default (top-right, bottom-left, and bottom-right, respectively)."
```
```{r, fig.cap= derivatives, fig.align="center"}
p1 <- ggdraw() + draw_image("../../Auxiliar/Pictures/function.jpg", scale = 0.9)
p2 <- ggdraw() + draw_image("../../Auxiliar/Pictures/first_derivative.jpg", scale = 0.9)
p3 <- ggdraw() + draw_image("../../Auxiliar/Pictures/second_derivative.jpg", scale = 0.9)
p4 <- ggdraw() + draw_image("../../Auxiliar/Pictures/third_derivative.jpg", scale = 0.9)

plot_grid(p1,p2,p3,p4, ncol = 2)  
```
```{r}
densities = "\\label{fig:densities} The figure plots probability of default histograms by portfolio for all banks and years. The last bin contains all observations with PD higher or equal to 25%. Defaulted exposures are excluded."
```

```{r, fig.height = 6, fig.width = 6, results="hide",fig.cap= densities, include=TRUE}

PD_summary <-  pillar3.data %>%
  select(name, PD, EAD, Portfolio_2, Portfolio_1) %>%
  filter(EAD>0, PD<1, Portfolio_2 %in% c("Banks", "Corporate", "Sovereign") | Portfolio_1 %in% c("Retail"))%>%
  na.omit(cols=c("EAD", "Portfolio")) %>%
  mutate(EAD = round(EAD/10))%>%
  expandRows("EAD")%>%
  mutate(Portfolio = ifelse(Portfolio_1 == "Retail", "Retail", as.character(Portfolio_2)),
         low25PD = ifelse(PD<0.25,1,0),
         low5PD = ifelse(PD<0.05,1,0)) %>%
           group_by(Portfolio) %>%
  dplyr::summarize(mean = round(100*mean(PD),2),
            sd   = round(100*sd(PD),2),
            p25 = round(100*sum(low25PD)/n(),2),
            p5 = round(100*sum(low5PD)/n(),2)) %>%
  mutate(lab = paste("Mean = ", mean, "% \n St. Dev. = ", sd, "% \n P(PD<25%) = ",p25,"% \n P(PD<5%) = ",p5, "%", sep = ""))

 pillar3.data %>%
 select(name, PD, EAD, Portfolio_2, Portfolio_1) %>%
  filter(EAD>0, PD<1, Portfolio_2 %in% c("Banks", "Corporate", "Sovereign") | Portfolio_1 %in% c("Retail"))%>%
  mutate(Portfolio = ifelse(Portfolio_1 == "Retail", "Retail", as.character(Portfolio_2)),
         EAD = round(EAD/10),
         PD = ifelse(PD>0.25, 0.25, PD)) %>%
  na.omit(cols=c("EAD", "Portfolio")) %>%
  expandRows("EAD")%>%
  ggplot(mapping = aes(x = PD)) +
    geom_histogram(aes(y=..density..), bins = 50)+
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey() +
  facet_wrap(~ Portfolio) +
  geom_text(data = PD_summary, aes(label = lab), x = 0.15, y = 150)

```

\newpage
\section{}  \label{sec:appendixb}

\setcounter{table}{0}
\renewcommand{\thetable}{B\arabic{table}}

\begin{table}[!htbp] \centering  
\caption{Summary Statistics}  
\label{tab:summary}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{13.5cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
\end{tabular}

```{r summary1,  fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
table <-capture.output({bank.data.win5 %>%
        filter(year>2007) %>% 
  select(
  log.RWA.r.Total_gr,
  log.RWA.r.Retail_gr,
  log.RWA.r.Wholesale_gr, 
  log.RWA.r.Corporate_gr, 
  log.RWA.r.Sovereign_gr, 
  log.RWA.r.Banks_gr,
  log.RWA.r.Total, 
  log.RWA.r.Retail,
  log.RWA.r.Wholesale, 
  log.RWA.r.Corporate, 
  log.RWA.r.Sovereign, 
  log.RWA.r.Banks,
  log.RWA.s.Total_gr,
  log.RWA.s.Retail_gr,
  log.RWA.s.Wholesale_gr, 
  log.RWA.s.Corporate_gr, 
  log.RWA.s.Sovereign_gr, 
  log.RWA.s.Banks_gr,
  RW.s.Total_gr,
  RW.s.Retail_gr,
  RW.s.Wholesale_gr, 
  RW.s.Corporate_gr, 
  RW.s.Sovereign_gr, 
  RW.s.Banks_gr,
  gini.PD.Total_gr,
  gini.PD.Retail_gr,
  gini.PD.Wholesale_gr, 
  gini.PD.Corporate_gr, 
  gini.PD.Sovereign_gr, 
  gini.PD.Banks_gr,
  log.CAR_gr,
  log.K_gr, 
  log.RWA.IRB_gr,
  log.RWA.hat.Total_gr,
  log.RWA.SA_gr,
  RW.SA_gr, 
  q.SA_gr, 
  q.Wholesale_gr, 
  q.Retail_gr) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
          label = "tab:summary",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = 
            c("\\hspace{0.0cm}$\\Delta \\text{Log RWA}^s$ total",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^s$ retail",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^s$ wholesale",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ corporate",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ sovereign", 
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ banks",
              "\\hspace{0.0cm}$ \\text{Log RWA}^s$ total",
              "\\hspace{0.5cm}$ \\text{Log RWA}^s$ retail",
              "\\hspace{0.5cm}$ \\text{Log RWA}^s$ wholesale",
              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ corporate",
              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ sovereign", 
              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ banks",
              "\\hspace{0.0cm}$\\Delta \\text{Log RWA}^{\\$}$ total",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ retail",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ wholesale",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ corporate",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ sovereign", 
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ banks",
              "\\hspace{0.0cm}$\\Delta \\text{RW}^s$ total",
              "\\hspace{0.5cm}$\\Delta \\text{RW}^s$ retail",
              "\\hspace{0.5cm}$\\Delta \\text{RW}^s$ wholesale",
              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ corporate",
              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ sovereign", 
              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ banks",
              "\\hspace{0.0cm}$\\Delta$ Log Gini total",
              "\\hspace{0.5cm}$\\Delta$ Log Gini retail",
              "\\hspace{0.5cm}$\\Delta$ Log Gini wholesale",
              "\\hspace{1.0cm}$\\Delta$ Log Gini corporate",
              "\\hspace{1.0cm}$\\Delta$ Log Gini sovereign", 
              "\\hspace{1.0cm}$\\Delta$ Log Gini banks",
              "$\\Delta \\text{Log capital ratio}$",
              "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\text{IRB}}$",
              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\text{SA}}$",
              "\\hspace{0.5cm}$\\Delta \\text{RW}^{\\text{SA}}$",
              "$\\Delta q^{\\text{SA}}$",
              "$\\Delta q^{\\text{IRB, Wholesale}}$",
              "$\\Delta q^{\\text{IRB, Retail}}$"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          summary = TRUE,
          digits = 3,
          align = TRUE,
          float = FALSE)})

table <- c(table[1:8],"Panel A: Main dependent variables \\\\",table[-(1:8)])

table <- c(table[1:21],"Panel B: Alternative measures \\\\",table[-(1:21)])
table <- c(table[1:21]," \\\\[-1.8ex]",table[-(1:21)])

table <- c(table[1:41],"Panel C: Other dependent variables \\\\",table[-(1:41)])
table <- c(table[1:41],"\\\\[-1.8ex]",table[-(1:41)])

table <- gsub("}lD{.}{.}{-3}",  "}p{0.35\\textwidth}D{.}{.}{4.1}", table, fixed = T)

#table <- gsub("{-3}", "{3.4}", table, fixed = T)

cat(table)

```

\end{table}

\begin{table}[!htbp] \centering
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{\parbox[c]{13cm}{\centering \tablename\ \thetable{} -- continued from previous page}}\\
\end{tabular}\\
\footnotesize

```{r summary3, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
table <- capture.output({bank.data.win5 %>%
                         filter(year>2007) %>% 
  select(
  mean.PD.Total_gr, 
  mean.PD.Retail_gr,
  mean.PD.Wholesale_gr, 
  mean.PD.Corporate_gr, 
  mean.PD.Sovereign_gr, 
  mean.PD.Banks_gr,
  var.PD.Total_gr, 
  var.PD.Retail_gr,
  var.PD.Wholesale_gr, 
  var.PD.Corporate_gr, 
  var.PD.Sovereign_gr, 
  var.PD.Banks_gr,
  skew.PD.Total_gr,
  skew.PD.Retail_gr,
  skew.PD.Wholesale_gr, 
  skew.PD.Corporate_gr, 
  skew.PD.Sovereign_gr, 
  skew.PD.Banks_gr,
  log.asset_gr,
  NPL.ratio_gr,
  log.asset,
  NPL.ratio,
  deposit.ratio,
  NII.ratio,
  Zscore,
  ROE,
  ROA,
  log.gdp_gr,
  credit_gdp_gr,
  log.gdp,
  credit_gdp,
  cap_reg,
  Sup_Power) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
          label = "tab:summary",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = 
            c("$\\Delta$ Log PD mean total",
              "\\hspace{0.5cm}$\\Delta$ Log PD mean retail",
              "\\hspace{0.5cm}$\\Delta$ Log PD mean wholesale",
              "\\hspace{1.0cm}$\\Delta$ Log PD mean corporate",
              "\\hspace{1.0cm}$\\Delta$ Log PD mean sovereign", 
              "\\hspace{1.0cm}$\\Delta$ Log PD mean banks", 
              "$\\Delta$ Log PD variance total",
              "\\hspace{0.5cm}$\\Delta$ Log PD variance retail",
              "\\hspace{0.5cm}$\\Delta$ Log PD variance wholesale",
              "\\hspace{1.0cm}$\\Delta$ Log PD variance corporate",
              "\\hspace{1.0cm}$\\Delta$ Log PD variance sovereign", 
              "\\hspace{1.0cm}$\\Delta$ Log PD variance banks", 
              "$\\Delta$ Log PD skewness total", 
              "\\hspace{0.5cm}$\\Delta$ Log PD skewness retail",
              "\\hspace{0.5cm}$\\Delta$ Log PD skewness wholesale",
              "\\hspace{1.0cm}$\\Delta$ Log PD skewness corporate",
              "\\hspace{1.0cm}$\\Delta$ Log PD skewness sovereign", 
              "\\hspace{1.0cm}$\\Delta$ Log PD skewness banks", 
              "$\\Delta \\text{Log total assets}$",
              "$\\Delta \\text{NPL / loans}$",
              "Log total assets",
              "NPL / loans",
              "Deposit / total assets",
              "NII / oper. rev.",
              "Z-score",
              "ROE",
              "ROA",
               "$\\Delta \\text{Log GDP}$",
              "$\\Delta \\text{Credit / GDP}$",
              "Log GDP",
              "Credit / GDP",
              "Capital stringency",
              "Official supervisory power"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          digits = 3,
          summary = TRUE,
          align = TRUE,
          float = FALSE)})

table <- c(table[1:8],"Panel D: PD distribution moments \\\\",table[-(1:8)])

table <- c(table[1:27],"Panel E: Other bank-level variables \\\\",table[-(1:27)])
table <- c(table[1:27]," \\\\[-1.8ex]",table[-(1:27)])

table <- c(table[1:39],"Panel F: Country-level variables \\\\",table[-(1:39)])
table <- c(table[1:39]," \\\\[-1.8ex]",table[-(1:39)])

table <- gsub("}lD{.}{.}{-3}",  "}p{0.35\\textwidth}D{.}{.}{4.1}", table, fixed = T)

cat(table)
```

\end{table}

\newpage
\section{}  \label{sec:appendixc}

\setcounter{table}{0}
\renewcommand{\thetable}{C\arabic{table}}

```{r reg_extension, echo=FALSE, include=FALSE}

# Variables in the model 
# Independent variables
aux.1 <- c("log.gdp_gr",
           "bank_credit_gdp_gr",
           "log.gdp",
           "cap_reg",
           "Sup_Power",
           "log.asset_gr",
           "NPL.ratio_gr",
           "log.asset",
           "NPL.ratio",
           "deposit.ratio",
           "NII.ratio")
ind.vars.temp  <- paste(aux.1, collapse=" + ")
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$",
              "$\\Delta\\text{Credit / GDP}_{i,t}$",
              "$\\text{Log GDP}_{i,t}$",
              "$\\text{Capital stringency}_{i,t}$", 
              "$\\text{Official supervisory power}_{i,t}$",
              "$\\Delta\\text{Log total assets}_{i,t}$",
              "$\\Delta\\text{NPL / loans}_{i,t}$",
              "$\\text{Log total assets}_{i,t}$",
              "$\\text{NPL / loans}_{i,t}$",
              "$\\text{Deposit / total assets}_{i,t}$",
              "$\\text{NII / oper. rev.}_{i,t}$")

# Model formulas
formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars.temp)," | year + bvdid | 0 | Country")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win5)
})

```

```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}

stargazer(reg.list,
          title = "Capital Savings and the Business Cycle: Control Variables",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^{s,j}_{i,t} =\\beta \\Delta\\text{Log GDP}_{i,t} + \\gamma \\Delta  X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$. In all regressions the measure of \textit{Capital savings} follows equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. $X_{i,t}$ is a vector of independent variables. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:extension",
          style            = "qje",
          column.sep.width = "-3pt",
          align            = TRUE)
```

\begin{table}[!htbp] \centering  
\caption{Capital Savings and the Business Cycle: Robustness to Alternative Samples}  
\label{tab:robustness}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Log RWA}_{i,t}^{s,j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$. In all regressions the measure of \textit{Capital savings} follows equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}\\
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: Balance panel} \\
\end{tabular}

```{r reg_constant, echo=FALSE}

reg.list <- list()
# Run the models
for(i in dep.vars){
formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country"))
 data = filter(bank.data.win5,
               get(paste0("log.RWA.r"  , i, "_gr_n"))>9)%>%
   felm(data = ., formula = formula.text ) -> reg.list[[i]]
  }

```

```{r table_constant, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```

\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\\
\multicolumn{6}{c}{Panel B: Sample winsorized at 1\%}
\end{tabular}

```{r reg_winsorized, echo=FALSE}

# Run the models
formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country"))
formulas <- lapply(formula.text, formula)
reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win1)
})

```

```{r table_winsorized, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```

\end{table}